---
title: "spatial data"
format: html
editor: visual
---

```{r}
knb_url <- "https://dev.nceas.ucsb.edu/knb/d1/mn/v2/object/urn%3Auuid%3Aaceaecb2-1ce0-4d41-a839-d3607d32bb58"

download.file(url = knb_url, destfile = 'shapefile_demo_data.zip')

unzip('shapefile_demo_data.zip', exdir = 'data')

file.remove('shapefile_demo_data.zip')
```

```{r}
library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggspatial)
library(dplyr)
```

```{r}
# read in shapefile using read_sf()
ak_regions <- read_sf("data/ak_regions_simp.shp")
# quick plot
plot(ak_regions)
```

```{r}
class(ak_regions)

head(ak_regions)

```

```{r}

glimpse(ak_regions)
```

```{r}
st_crs(ak_regions)
```

```{r}
ak_regions_3338 <- ak_regions %>%
    st_transform(crs = 3338)

st_crs(ak_regions_3338)

plot(ak_regions_3338)
```

-   Tidyverse

```{r}
# returns the names of all the columns in dataset
colnames(ak_regions_3338)

ak_regions_3338 %>%
    select(region)
```

```{r}
unique(ak_regions_3338$region)
```

```{r}
ak_regions_3338 %>%
    filter(region == "Southeast")
```

-   Spatial Joins

```{r}
# read in population data
pop <- read_csv("data/alaska_population.csv")
head(pop)
class(pop)
```

-   transformation pop into spatial object (to geometry from table)

```{r}
pop_4326 <- st_as_sf(pop,
                     coords = c('lng', 'lat'),
                     crs = 4326,
                     remove = F) #dont remove column

head(pop_4326)
class(pop_4326)
```

-   join population data with regions

```{r}
#pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within)# different projections, error


```

-transform CRS of pop data

```{r}
pop_3338 <- st_transform(pop_4326, crs = 3338)
st_crs(pop_3338)
```

```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)
head(pop_joined)
class(pop_joined)
```

```{r}
pop_region <- pop_joined %>%
    as.data.frame() %>%
    group_by(region) %>%
    summarise(total_pop = sum(population))

head(pop_region)
  
```

```{r}
pop_region_3338 <- left_join(ak_regions_3338,
                             pop_region,
                             by = "region")
head(pop_region_3338)

plot(pop_region_3338["total_pop"])
```

```{r}
pop_mgmt_3338 <- pop_region_3338 %>%
    group_by(mgmt_area) %>%
    summarize(total_pop = sum(total_pop))

plot(pop_mgmt_3338["total_pop"])
```

-save spatial object

```{r}
write_sf(pop_region_3338, "data/ak_regions_population.shp")
```

-vuzualise

```{r}
ggplot(pop_region_3338) +
    geom_sf(aes(fill = total_pop)) +
    labs(fill = "Total Population") +
    scale_fill_continuous(low = "khaki",
                          high =  "firebrick",
                          labels = comma) +
    theme_bw()
```

```{r}
rivers_3338 <- read_sf("data/ak_rivers_simp.shp")
st_crs(rivers_3338)
```

```{r}
ggplot() +
    geom_sf(data = pop_region_3338, 
            aes(fill = total_pop)) +
    geom_sf(data = pop_3338, 
            size = 0.5) +
    geom_sf(data = rivers_3338,
            aes(linewidth = StrOrder)) +
    scale_linewidth(range = c(0.05, 0.5), 
                    guide = "none") +
    labs(title = "Total Population by Alaska Region",
         fill = "Total Population") +
    scale_fill_continuous(low = "khaki",
                          high =  "firebrick",
                          labels = comma) +
    theme_bw() 
```

### Incorporate base maps into static maps using `ggspatial`

```{r}
pop_3857 <- pop_3338 %>% 
  st_transform(crs = 3857)

ggplot(data = pop_3857) +
    ggspatial::annotation_map_tile(type = "osm", zoom = 4) + # higher zoom values are more detailed 
    geom_sf(aes(color = population),
            fill = NA) +
    scale_color_continuous(low = "darkkhaki",
                           high =  "firebrick",
                           labels = comma)
```

-leaflet

```{r}
epsg3338 <- leaflet::leafletCRS(
    crsClass = "L.Proj.CRS",
    code = "EPSG:3338",
    proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
    resolutions = 2 ^ (16:7)
)

st_crs(pop_region_3338)
```

```{r}
pop_region_4326 <- pop_region_3338 %>% 
    st_transform(crs = 4326)
```

```{r}
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
    addPolygons(data = pop_region_4326,
                fillColor = "gray",
                weight = 1)

m
```

```{r}
pal <- colorNumeric(palette = "Blues", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
    addPolygons(
        data = pop_region_4326,
        fillColor = ~ pal(total_pop),
        weight = 1,
        color = "black",
        fillOpacity = 1,
        label = ~ region
    ) %>%
    addLegend(
        position = "bottomleft",
        pal = pal,
        values = range(pop_region_4326$total_pop),
        title = "Total Population"
    )

m
```

```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
    addPolygons(
        data = pop_region_4326,
        fillColor = ~ pal(total_pop),
        weight = 1,
        color = "black",
        fillOpacity = 1
    ) %>%
    addCircleMarkers(
        data = pop_4326,
        lat = ~ lat,
        lng = ~ lng,
        radius = ~ log(population / 500),
        # arbitrary scaling
        fillColor = "gray",
        fillOpacity = 1,
        weight = 0.25,
        color = "black",
        label = ~ paste0(pop_4326$city, ", population ", comma(pop_4326$population))
    ) %>%
    addLegend(
        position = "bottomleft",
        pal = pal,
        values = range(pop_region_4326$total_pop),
        title = "Total Population"
    )

m

